What should I notice? Using Alogithmic Information Theory to evaluate the memorability of events in smart homes

### 2021-11-03 new_paper.tex
Fixing following Ada's comments pt2

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
\usepackage{algorithm2e}

@@

%=================================================================
% Full title of the paper (Capitalized)
\Title{What should I notice? Evaluating the memorability of events for abductive inference.
}

@@

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{ When confronted with an unprecedented situation, humans typically show good
performance in quickly identifying noticeable past events and proposing them
as possible causal hypotheses. This kind of abductive inference is widely
overlooked in modern AI approaches which rely on massive datasets to learn
associative patterns. Our proposal is to formalize and compute a ``memorability''
score over a set of events recorded from a cyber-physical system.
This score can then be used to select relevant information to be
remembered and to propose causal hypotheses in unusual situations on demand.
The approach is meant to be complementary to more traditional
learning-focused abduction techniques. We provide a theoretical framework based on Algorithmic Information Theory to define the notion of memorability of events. Then, we implement our approaches in two examples based on smart-home scenario to validate our approach in toy cases and illustrate how it could be applied.}

% Keywords
\keyword{Kolmogorov Complexity; Algorithmic Information Theory; Simplicity; Abduction; Surprise}

@@

\section{Introduction}

@@

In this example, there are mainly three possibilities to come to the
conclusion. (1) If the user knows how the smart living-room system works, if she
knows the underlying rules or parameters, she may use this causal
knowledge to perform abduction. (2) If she has no knowledge
about the system but made several observation of the same behavior, she may examine past
correlations and figure out that turning on the TV set often leads the blinds to
close and the lights to dim. (3) If there are no previous occurrences of the event (e.g. it is the first time she turns on the TV
in the living-room), she may still be able to
suspect that the TV is a possible cause for the observed event, just because it appears to her as a memorable recent event (as it is its first occurrence). This example
suggests that human beings are able to use distinct methods to perform abductive tasks
and infer new knowledge, depending on the situation. While the first two mechanisms can be
automated using knowledge bases and statistical methods,the third approach, which can be used without any knowledge of the occurring phenomenon or past occurrences. Instead, it relies on the observation that some events appear more ``memorable'' than others and, as such, can be considered as possible hypotheses if need be.

@@

% ? What to do with the "intuition" thing? Keep it? Leave it?
To address this issue, we started from the following intuition: while all
events, regardless of their characteristics or nature, can be uniquely
described using a combination of quantitative or qualitative qualifiers, the most memorable ones are
likely to require less words to be described. ``Last year's hottest day'' is simpler than ``the 182\textsuperscript{nd} day 7 years ago''. How can we quantify this relative simplicity? We could evaluate the complexity of each
description, taking into account both the complexity of the concept words (a date
of occurrence, a temperature ranking), and of the arguments (1st hottest,
182 and 7). The resulting values define the  \emph{description complexity}
of events. Our intuition is that memorable events require simpler and less numerous qualifiers to be
unambiguously described than unremarkable ones.
% ? Response to Ada's remark: added a thing in the 1st paragraph to show what we want ot do with the memorable events.

For machines to implement and compute description complexity, we need a formal framework and computation methods that are in line with human intuition.  Algorithmic Information Theory (AIT) appears to be such a framework, as it is consistent with the human perception of complexity\cite{li_introduction_2008,dessalles2011coincidences,delahaye_numerical_2012}. Our method is as follows: we
consider events as being elements stored in what we call \emph{base memory}. To
reproduce the language features applicable to events, we use \emph
{predicates}, i.e. functions assigning a boolean value to events. For instance, the predicate $\mathtt{date}
(\cdot, 1\_year)$ is \emph{true} of events that occurred last year. Selecting
%%%%%%%%%%%%%%%%%%%%%% ! MOVED FIGURE HERE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[hb]
    \centering
    \includegraphics[width=.95\linewidth]{figures/filters}
    \caption{Retrieving an event through successive predicative
        filters. From the base memory (yellow), successive filters select events satisfying the associated predicate (gray arrows). For example,
        filter $f_1$ selects events from last year, i.e. which satisfy the
        predicate $\mathtt{date}(event, 1\_year)$. In this case, successively
        applying filters $f_1$, $f_2$ and $f_3$ yields a unique event $e_h$, last
        year's hottest day. The complexity of this event can then be upper-bounded
        by the complexity of the three filters as they give an unambiguous way
        to describe the event within the memory.}
    \label{fig:filters}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%! FIGURE END %%%%%%%%%%%%%%%%%%%%%%%%%%%%
all events, from the memory, that satisfy a given predicate corresponds to
a \emph{filter} operation. It generates another memory that is a subset of the previous one.
The filtering operation can then be repeated, selecting
fewer events at each iteration, until a singleton memory is reached. This means that the sequence
of predicates could unambiguously \emph{retrieve} the unique remaining event.
The description complexity of this event can thus be upper-bounded by the
number of bits required to describe the filters used in the retrieval
process. Figure \ref{fig:filters} illustrates this process for the event: ``last year's hottest day''.

% * Paragraph re-written
We will first briefly introduce
some basic notions of Algorithmic Information Theory in subsection~\ref{sec:theory}. We then expose our contribution by presenting in subsection~\ref{sec:computing} our formal definition of memorability and related notions. Then, we present an implementation example of these definitions with two smart-home examples in section~\ref{sec:results}. The results of these experiments are then presented and discussed. Finally we explore other related works in subsection \ref{sec:related} and explore possible extensions of our work in subsection \ref{sec:future}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

\subsection{Theoretical Background}
\label{sec:theory}
Kolmogorov complexity formally quantifies the amount of information required
for the computation of a finite binary string\footnote{Though the definition
    holds for some infinite binary strings (think of the representation of the
    decimals of $\pi$), we restrict ourselves here to finite strings.} (or
any object represented by a finite binary
string)\cite{kolmogorov_three_1965,li_introduction_2008}. The complexity $K(s)$ of a (finite) binary string is the length in bits $L(p)$ of the shortest program $p$
which, if given as input to a universal Turing Machine $U$, outputs $s$.

@@

As \cite{dessalles2011coincidences} suggests a link between the unexpectedness and cognitive relevance, we propose to define the memorability of an event in a similar way.  
Hence, we define the memorability $M(e)$ of an event as the absolute difference between its description complexity $K_d(e)$ and its expected description complexity $K_{exp}(e)$:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{sec:results}
\subsection{Setup}

We design two different setups to test our approach. Both are inspired from smart home use cases.  This choice of
configuration is motivated by the challenges posed by smart homes for abductive
inference: i) as the number of connected devices increases, more events are
recorded, making the detection of memorable events more important; ii) smart
homes are prone to experience atypical situations, highly dependent on the context,
for which pre-established relations might fail to find good abduction
candidates. Our choice was also motivated by
the existence of previous work involving smart home simulations capable of quickly generating data from which we could extract events and test our methods.

@@

\subsubsection{Scenario 2}
We consider an experimental smart home setup, with various sensors, which we simulate over a period of time
To carry out the simulation, we built custom modules into the existing iCasa
smart home simulator\cite{lalanda_self-aware_2017}. ICasa
offers a simulation of autonomic systems that can handle internal communications,
the possible insertion of new components at runtime, or the deletion or modification of existing
components. We used a basic scenario consisting of a house with four rooms,
a single user, and an outdoor zone. All four rooms are equipped
with a temperature controller system that monitors and controls heaters (fig.
\ref{fig:view}).

\begin{figure}[ht]
    \centering
    \includegraphics[width=.8\linewidth]{figures/simulator}
    \caption{View of the simulator's web interface provided by iCasa. The four
        rooms are visible, with their equipment and the user.}
    \label{fig:view}
\end{figure}

@@

\begin{figure}[ht]
    \includegraphics[width=\linewidth]{figures/ts_example}
    \caption{Time series data from the simulation: outdoor temperature (blue) and
        controller temperature of a room (orange). To be used in our framework, these time series data are
        processed by a simple threshold-based event detection.}
    \label{fig:ts_example}
\end{figure}

The values observed from all devices and zones was sampled throughout
the simulation run. The resulting data (figure
\ref{fig:ts_example}) was then used as a basis for our experiments.
We then process the time series data to identify and characterize events.
Since the selection of events is not the focus point of our present work (see sec.
\ref{sec:related}), we perform event detection merely based on threshold and precomputed conditions.

@@

We implemented the computation of both the description complexity and the memorability score into a Python object, called the \texttt{SurpriseAbductionModule}. Apart from the base memory of events $M_0$, this module contains a set of predefined predicates $\Pi$ to characterize events. For instance, for scenario 2, the predicates we use are the following:
\begin{itemize}
    \item $\mathtt{label(e, k)}$: whether the event $e$ has the label ranked
          $k^{th}$ in the label list.
    \item $\mathtt{rank(e, r, a)}$: whether the event $e$ has the rank $r$
          along axis $a$.
    \item $\mathtt{day(e, k)}$: whether the event $e$ occurred $k$ days ago.
          % \item $\mathtt{randomChoice(e, k)}$: whether the event $e$ id matches the program $k$, which can be regarded as a random seed.
    \item $\mathtt{month(e, k)}$: whether the event $e$ occured $k$ months ago.
    \item $\mathtt{location(e, k)}$: whether the event $e$ occurred in zone
          $k$.
\end{itemize}

Similar predicates were used for scenario 1, with the notable exception of the addition of a \texttt{RandomPick} predicate, which is true if the id of the event matches a random program seed $k$. Using this predicate and the associated filter in only one of the two illustrative examples shows the importance of allowing random selection in the complexity computation. Furthermore, it could be argued that in some case, random selection of an event is not possible.

The description length $L_{\pi, k}$ of using a predicate is computed as follows: since the set of predicates is finite and known, $L(\pi) = \log2(|\Pi|)$ are enough to describe the predicate concept $\pi$\footnote{This approach gives an equal complexity to all predicate concepts. While this could be argued when using many concepts, as humans do, we deemed better, for our small examples, to consider all concepts as equal.}. To encode the argument $k$ of the predicate, we used the prefix-free Elias delta code\cite{elias_universal_1975}, which requires $L(k) = \log2(k) + 2 \log2(\log2(k)+1) + 1]$ bits. The total cost of describing $\pi_{k}$ therefore is

\begin{equation}
    \L(\pi, k) = \log2(|\Pi|) + \log2(k) + 2 \log2(\log2(k) + 1) + 1
\end{equation}

@@

@@

\subsubsection{The ``TV'' scenario:}
\paragraph{Memorability}
The computation of the description complexity measure for the 2400 events recorded in this scenario took around 90 seconds using an i7-8600u laptop. The resulting complexities, in Figure \ref{fig:scenar1_complexity}, show that TV events appear more simple than luminosity events. In fact, for most luminosity events, the complexity corresponds to the random choice complexity, 25 bits in this example. This pattern comes from the number of luminosity events per day, 24, is too high to consider the retrieval path $day,  randomPick$ simpler than $randomPick$ from the entire memory, except for the most recent days, which appear simpler (on the right-hand side of Figure \ref{fig:scenar1_complexity})
\begin{figure}
    \centering
    \includegraphics[width=.7\linewidth]{figures/memorability_scenar_1.png}
    \caption{Description complexity for events recorded in the TV scenario. Luminosity events are shown in blue, TV events in green}
    \label{fig:scenar1_complexity}
\end{figure}

@@

\begin{table}
    \centering
    \begin{tabular}{l|c|c}
        Event Id & Description                   & Relative memorability (bits) \\
        \hline
        2513     & Use of smart TV               & 16.76                        \\
        2427     & Last use of the old TV        & 14.81                        \\
        2411     & Second-last use of the old TV & 11.21
    \end{tabular}
    \caption{Output of the memorability-based abduction module: top 3 events for the relative memorability metrics.}
    \label{tab:abduction_res}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

\subsection{Related Works}
\label{sec:related}
Our work is intended to be integrated into larger-scale frameworks to monitor
and detect events in complex environment such as smart homes. In these works, the
approach to smart homes is often regarded as self-organizing systems \cite{kramer_rigorous_2009,kounev_notion_2017}. As such, they present capacities of adaptation to new goals,
new components, new environment. A commonly used approach is the principle of
autonomic system, which minimizes user's intervention for management of the
system \cite{kounev_notion_2017,kephart_vision_2003}.

@@

\subsection{Perspectives}
\label{sec:future}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
The introduction of description complexity and the subsequent memorability measure allow to compare events of various types with different characteristics. As such, it provides a formal and generic framework to discriminate peculiar events and outliers.

The computation being based on predicates, the memorability score is entirely dependent on their definitions. This can be seen as a caveat or a feature of our approach: it makes the measure subjective while enabling the consideration of user's preferences in the definition and choice of predicates. This consideration can be the topic of future research and open up potential applications.

>>>>

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
\usepackage{algorithm2e}
\usepackage{caption}
\usepackage{subcaption}

@@

%=================================================================
% Full title of the paper (Capitalized)
\Title{What should I notice? Using Alogithmic Information Theory to evaluate the memorability of events in smart homes
}

@@

\abstract{
With the increasing number of connected devices, complex systems such as smart homes record a multitude of events of various types, magnitude and characteristics. Faced with this variety and number, current systems struggle to identify which events can be considered more memorable than others. On the other hand, human beings are able, without knowledge of the system's inner working or large previous datasets, to quickly categorize some events as being more ``memorable'' than others. Having this ability would allow the system to identify and summarize a situation to the user or use the most memorable events as possible hypotheses in an abducitve inference process. Our proposal is to use Algorithmic Information Theory to formally define a ``memorability'' score based on the concept of retrieving events by using predicative filters. After providing a theoretical framework, we provide smart-home examples to illustrate how our approach could be implemented in practive and qualitatively compare our ``memorability'' score with human perception.
}

% When confronted with an unprecedented situation, humans typically show good
% performance in quickly identifying noticeable past events and proposing them
% as possible causal hypotheses. This kind of abductive inference is widely
% overlooked in modern AI approaches which rely on massive datasets to learn
% associative patterns. Our proposal is to formalize and compute a ``memorability''
% score over a set of events recorded from a cyber-physical system.
% This score can then be used to select relevant information to be
% remembered and to propose causal hypotheses in unusual situations on demand.
% The approach is meant to be complementary to more traditional
% learning-focused abduction techniques. We provide a theoretical framework based on Algorithmic Information Theory to define the notion of memorability of events. Then, we implement our approaches in two examples based on smart-home scenario to validate our approach in toy cases and illustrate how it could be applied.}

\keyword{Kolmogorov Complexity; Algorithmic Information Theory; Simplicity; Abduction; Surprise}

@@

\section{Introduction}
\label{sec:introduction}

@@

In this example, there are mainly three possibilities to come to the
conclusion. (1) If the user knows how the smart living-room system works, if she
knows the underlying rules or parameters, she may use this causal
knowledge to perform abduction. (2) If she has no knowledge
about the system but made several observation of the same behavior, she may examine past
correlations and figure out that turning on the TV set often leads the blinds to
close and the lights to dim. (3) If there are no previous occurrences of the event (e.g. it is the first time she turns on the TV
in the living-room), she may still be able to
suspect that the TV is a possible cause for the observed event, just because it appears to her as a memorable recent event (as it is its first occurrence). This example
suggests that human beings are able to use distinct methods to perform abductive tasks
and infer new knowledge, depending on the situation. While the first two mechanisms can be
automated using knowledge bases and statistical methods,the third approach, which can be used without any knowledge of the occurring phenomenon or past occurrences. Instead, it relies on the observation that some events appear more ``memorable'' than others and, as such, can be considered as possible hypotheses if need be.\cite{magnani_abduction_2011}

@@

% ? What to do with the "intuition" thing? Keep it? Leave it?
To address this issue, we started from the following supposition: while all
events, regardless of their characteristics or nature, can be uniquely
described using a combination of quantitative or qualitative qualifiers, the most memorable ones are
likely to require less words to be described. This supposition appears to be in line with some observations: for instance, a corrolation has been found between word frequency and length\cite{strauss_word_2007}. Thus, ``last year's hottest day'' is simpler than ``the 182\textsuperscript{nd} day 7 years ago''. How can we quantify this relative simplicity? We could evaluate the complexity of each
description, taking into account both the complexity of the concept words (a date
of occurrence, a temperature ranking), and of the arguments (1st hottest,
182 and 7). The resulting values define the  \emph{description complexity}
of events. Our intuition is that memorable events require simpler and less numerous qualifiers to be
unambiguously described than unremarkable ones.
% ? Response to Ada's remark: added a thing in the 1st paragraph to show what we want ot do with the memorable events.

For machines to implement and compute description complexity, we need a formal framework and computation methods that incorporate the aforementionned process. Algorithmic Information Theory (AIT) appears to be such a framework, as it is consistent with the human perception of complexity\cite{li_introduction_2008,dessalles2011coincidences,delahaye_numerical_2012}. Our method is as follows: we
consider events as being elements stored in what we call \emph{base memory}. To
reproduce the language features applicable to events, we use \emph
{predicates}, i.e. functions assigning a boolean value to events. For instance, the predicate $\mathtt{date}
(\cdot, 1\_year)$ is \emph{true} of events that occurred last year. Selecting
\begin{figure}[hb]
    \centering
    \includegraphics[width=.95\linewidth]{figures/filters}
    \caption{Retrieving an event through successive predicative
        filters. From the base memory (yellow), successive filters select events satisfying the associated predicate (gray arrows). For example,
        filter $f_1$ selects events from last year, i.e. which satisfy the
        predicate $\mathtt{date}(event, 1\_year)$. In this case, successively
        applying filters $f_1$, $f_2$ and $f_3$ yields a unique event $e_h$, last
        year's hottest day. The complexity of this event can then be upper-bounded
        by the complexity of the three filters as they give an unambiguous way
        to describe the event within the memory.}
    \label{fig:filters}
\end{figure}
all events, from the memory, that satisfy a given predicate corresponds to
a \emph{filter} operation. It generates another memory that is a subset of the previous one.
The filtering operation can then be repeated, selecting
fewer events at each iteration, until a singleton memory is reached. This means that the sequence
of predicates could unambiguously \emph{retrieve} the unique remaining event.
The description complexity of this event can thus be upper-bounded by the
number of bits required to describe the filters used in the retrieval
process. Figure \ref{fig:filters} illustrates this process for the event: ``last year's hottest day''.

% * Paragraph re-written 
The rest of this article is organized as follows.
We will first briefly introduce
some basic notions of Algorithmic Information Theory in subsection~\ref{sec:theory}. We then expose our contribution by presenting in subsection~\ref{sec:computing} our formal definition of memorability and related notions. Then, we present an implementation example of these definitions with two smart-home examples in section~\ref{sec:results}. The results of these experiments are then presented and discussed. Finally we explore other related works in subsection \ref{sec:related} and explore possible extensions of our work in subsection \ref{sec:future}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theoretical Framework}
\label{sec:framework}

\subsection{Background}
\label{sec:theory}
Kolmogorov complexity formally quantifies the amount of information required
for the computation of a finite binary string\footnote{Though the definition
    holds for some infinite binary strings (think of the representation of the
    decimals of $\pi$), we restrict ourselves here to finite strings.} (or
any object represented by a finite binary
string)\cite{kolmogorov_three_1965,li_introduction_2008}. The complexity $K(s)$ of a (finite) binary string is the length in bits $L(p)$ of the shortest program $p$
which, if given as input to a universal Turing Machine $U$, outputs $s$.

@@

As \cite{dessalles2011coincidences} suggests a link between the unexpectedness and cognitive relevance, we propose to define the memorability of an event in a similar way. Since we want to use this score in applications, we need a definition that is well-defined and computable in practice.
Hence, we define the memorability $M(e)$ of an event as the absolute difference between its description complexity $K_d(e)$ and its expected description complexity $K_{exp}(e)$:

@@

\section{Experiments}
\label{sec:results}
\subsection{Setup}

We design two different setups to test our approach. Both are inspired from smart home use cases.  This choice of
configuration is motivated by the challenges posed by smart homes for abductive
inference: i) as the number of connected devices increases, more events are
recorded, making the detection of memorable events more important; ii) smart
homes are prone to experience atypical situations, highly dependent on the context,
for which pre-established relations might fail to find good abduction
candidates. Our choice was also motivated by
the existence of previous work\cite{lalanda_self-aware_2017} involving smart home simulations capable of quickly generating data from which we could extract events and test our methods.

@@

\subsubsection{Scenario 2}
We consider an experimental smart home setup, with various sensors, which we simulate over a period of time.
To carry out the simulation, we built custom modules into the existing iCasa
smart home simulator\cite{lalanda_self-aware_2017}. ICasa
offers a simulation of autonomic systems that can handle internal communications,
the possible insertion of new components at runtime, or the deletion or modification of existing
components. We used a basic scenario consisting of a house with four rooms,
a single user, and an outdoor zone. All four rooms are equipped
with a temperature control system that controlling heaters (see Figure.
\ref{fig:view}).

\begin{figure}[ht]
    \centering
    \includegraphics[width=.8\linewidth]{figures/icasa_interface.png}
    \caption{View of the simulator's web interface provided by iCasa. The four
        rooms are visible, with their equipment and the user.}
    \label{fig:view}
\end{figure}

@@

\begin{figure}[ht]
    \includegraphics[width=\linewidth]{figures/ts_example}
    \caption{Time series data from the simulation: outdoor temperature (red) and
        controller temperature of a room (blue). To be used in our framework, these time series data are
        processed by a simple threshold-based event detection.}
    \label{fig:ts_example}
\end{figure}

Values observed from all devices and zones were sampled throughout
the simulation. The resulting data (figure
\ref{fig:ts_example}) was then used as a basis for our experiments.
We then process the time series data to identify and characterize events.
Since how events are detected is not the focus point of our present work (see Section.
\ref{sec:related}), we perform event detection merely based on threshold and precomputed conditions.

@@

We implemented the computation of both the description complexity and the memorability score into a Python object, called the \texttt{SurpriseAbductionModule}. Apart from the base memory of events $M_0$, this module contains a set of predefined predicates $\Pi$ to characterize events. For instance, for scenario 2, the predicates we use are the following:
\begin{itemize}
    \item $\mathtt{label(e, k)}$: whether the event $e$ has the $k^{th}$ most frequent label (meaning that frequent labels are simpler to express than rare ones)
    \item $\mathtt{rank(e, r, a)}$: whether the event $e$ is ranked $r^{th}$ for characteristic $a$, where characteristics are encoded by their frequency (again, common characteristics are the simplest ones)
    \item $\mathtt{day(e, k)}$: whether the event $e$ occurred $k$ days ago.
          % \item $\mathtt{randomChoice(e, k)}$: whether the event $e$ id matches the program $k$, which can be regarded as a random seed.
    \item $\mathtt{month(e, k)}$: whether the event $e$ occured $k$ months ago.
    \item $\mathtt{location(e, k)}$: whether the event $e$ occurred in zone
          $k$.
\end{itemize}

% Similar predicates were used for scenario 1. However, to show the impact of allowing random selection of events, we added a \texttt{RandomPick} predicate, which is true if the id of the event matches a random program seed $k$. As it could be argued that in some cases, randomly selecting an event is not possible, we decided to implement this predicate in only one scenario.

% with the notable exception of the addition of a \texttt{RandomPick} predicate, which is true if the id of the event matches a random program seed $k$. Using this predicate and the associated filter in only one of the two illustrative examples shows the importance of allowing random selection in the complexity computation. Furthermore, it could be argued that in some case, random selection of an event is not possible.

The description length $L(\pi, k)$ of a predicate $\pi_k$ is computed as follows: since the set of predicates is finite and known, $L(\pi) = \log2(|\Pi|)$ are enough to describe the predicate concept $\pi$\footnote{This approach gives an equal complexity to all predicate concepts. While this could be argued when using many concepts, as humans do, we deemed better, for our small examples, to consider all concepts as equal.}. To encode the argument $k$ of the predicate, we used the widely used prefix-free Elias delta code\cite{elias_universal_1975}, which requires $L(k) = \log2(k) + 2 \log2(\log2(k)+1) + 1]$ bits. The total cost of describing $\pi_{k}$ therefore is

\begin{equation}
    \label{eq:pred_cost}
    L(\pi, k) = \log2(|\Pi|) + \log2(k) + 2 \log2(\log2(k) + 1) + 1
\end{equation}

@@

Our code is written in Python. Examples are presented in the form of Jupyter Notebooks, which allow to quickly reproduce our results and figures. All code is available on our Github: \url{https://github.com/EtienneHouze/memorability_code}.

@@

\subsubsection{The ``TV'' scenario:}
\paragraph{Memorability}
The computation of the description complexity measure for the 2400 events recorded in this scenario took around 90 seconds using an i7-8600u laptop. The resulting complexities, in Figure \ref{fig:scenar1_complexity}, show that TV events appear simpler than luminosity events. In fact, for most luminosity events, the complexity corresponds to the random choice complexity, 25 bits in this example. This is due to the frequency of luminosity events: given that there are 24 of them per day, considering the retrieval path $day(k), randomPick(24)$ costs more bits than $randomPick(N)$, where $N$ is the total number of events, with the notable exception of recent days, for which the value $k$ is low (see the right-hand side of Figure \ref{fig:scenar1_complexity}).

% TODO add some annotations to the figure?
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{figures/memorability_scenar_1.png}
    \caption{Description complexity for events recorded in the TV scenario. Luminosity events are shown in blue, TV events in green}
    \label{fig:scenar1_complexity}
\end{figure}

@@

% ! Response to Ada's comment: Indeed, the jump in memorability does not seem that high. However, we may consider that the smart TV event will remain memorable through time, as it will remain the first use of the device 'smart TV', whereas the two other events will become less memorable as time passes by... Maybe add some paragraph in the discussions section to talk about the phenomenon of considering recent events memorable?
\begin{table}
    \centering
    \begin{tabular}{l|c|c}
        Event Id & Description                   & Relative memorability (bits) \\
        \hline
        2513     & Use of smart TV               & 16.76                        \\
        2427     & Last use of the old TV        & 14.81                        \\
        2411     & Second-last use of the old TV & 11.21
    \end{tabular}
    \caption{Output of the memorability-based abduction module: top 3 events for the relative memorability metrics.}
    \label{tab:abduction_res}
\end{table}

@@

\subsection{Discussion: the subjectivity of predicates}
For individuals, the notion of memorability and complexity of events is highly subjective: the same event may appear usual for a person, while being exceptional for her neighbor. Our approach to memorbaility embraces this subjectivity, incorporating it through the definiton of predicates.

In Figure \ref{fig:3preds}, we show this subjectivity by comparing the analyses of the same base memory of events, generated using the same setup as for Example 1, using different three different sets of predicates. The resulting figures show different phenomena based on the predicates at hand by the module. In figure \ref{fig:pred_1}, only time and labels were recognized by predicates. As a result, the general trend of the curve is logarithmic, as events from $T$ days ago require $O(\log(T)))$ bits of information. Two main sequences of light events appear: as some light events were recorded simultaneously as TV events, they require the additional information of their label to be retreived. In Figure \ref{fig:pred_2}, we added predicates qualifying the light intensity, along with a day/night distinction. This added capacity isolated some light events as much simpler: they appear as a green horizontal line. These events are times when the light was low during day, or high during night. As such, they are outliers, and therefore require less inforamtion to be retrieved.

Finally, Figure \ref{fig:pred_3} comes from a module which has the ability to retrieve any event from a memory of size $N$, at the expense of $O(\log(N))$ bits of information. This added capacity has the immediate effect of setting a clear upper limit to the description complexity of items, since any item can be retrieved using $\l(N)$ bits (in this example, this limit is around 25 bits). While this kind of direct retrieval is trivial in computer science (one could use the memory address of any stored event), its correlation with human cognition is not obvious: can humans be considered to have this possibility of selecting any event from their entire memory, without restriction regarding their nature, their age, their magnitude? However, this highlights an interesting phenomenon: using this direct retrieval, the module makes no distinction, complexity-wise, between events past a given threshold. In other words, there is a generic "uninteresting" category of events, in which the module makes no distinction.

In addition to the number of available predicate concepts, one might also tweak their complexity. In our example scenarios, we used only a handful of predicates, hence we chose to assign all predicate concepts the same bit description length, $\log(|\Pi|)$ (see Equation. \ref{eq:pred_cost}). When using more predicate concepts, we may instead give different costs to some of them, to take into account the complexity difference between them: for instance, the generic time concept $\texttt{year}$ would likely require less bits than a predicate $\texttt{tun}$ (an old Mayan time unit, corresponding to roughly 360 days). Similar to the selection of predicates to be used, the complexity assigned to each predicate concept, that is, their encoding, denotes the subjectivity of the user.
\newpage

\begin{figure}[H]
    \centering
    \begin{subfigure}{\linewidth}
        \centering
        \includegraphics[width=.7\linewidth]{figures/preds_1.png}
        \caption{}
        \label{fig:pred_1}
    \end{subfigure}
    \begin{subfigure}{\linewidth}
        \centering
        \includegraphics[width=.7\linewidth]{figures/preds_2.png}
        \caption{}
        \label{fig:pred_2}
    \end{subfigure}
    \begin{subfigure}{\linewidth}
        \centering
        \includegraphics[width=.7\linewidth]{figures/preds_3.png}
        \caption{}
        \label{fig:pred_3}
    \end{subfigure}
    \caption{The same memory of events, analyzed through three different sets of predicates. Light events are in blue, TV events in salmon. Figure. \ref{fig:pred_1} uses only time-related predicates (days and hours ago), while Figure. \ref{fig:pred_2} adds label predicates alongside with "dark" and day/night predicates. Figure \ref{fig:pred_3} adds the possibility to select directly select any event from the memory of size $N$, at the cost of $l(N)$ bits.}
    \label{fig:3preds}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Works}
\label{sec:related}
Our work is intended to be integrated into larger-scale frameworks to monitor
and detect events in complex environment such as smart homes. In these works, the
approach to smart homes is often regarded as self-organizing systems \cite{kramer_rigorous_2009,kounev_notion_2017}. As such, they present capacities of adaptation to new goals,
new components, new environment. A commonly used approach is the principle of
autonomic system, which minimizes user's intervention for management of the
system \cite{kounev_notion_2017,kephart_vision_2003}.

@@

\section{Perspectives}
\label{sec:future}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
% TODO Re-write the conclusion!
The ability to identify some events as memorable is useful in the current context of computing, where connected devices record many events of different characteristics, magnitudes, types. This is particularly true in complex cyber-physical systems such as smart homes. We propose an approach to evaluate the memorability as a difference between the expected description complexity of an event and its actual value. With this definition, something is memorable if it requires much less or much more information to be described. To formalize this notion, we used the minimum Minimum Description Length principle exposed in Algorithmic Information Theory. By defining filter operations from predicates and successively applying these filters, we define formal ways of describing events. The length of the shortest of such way can be used as a measure of description complexity.

We provided an implementation algorithm to compute this measure on events and showed some of its results in two smart home examples. These application scenarios qualitatively show how our measure fares in comparision with the human notion of memorability, and how this measure can be used to propose relevant hypotheses in an abductive inference process without having further knowledge of the system.

Finally, we discussed the inherent subjectivity of our approach, embodied in the choice of available predicates and their encoding. We link our approach to existing works which use Algorithmic Information Theory principles to analyze time series data or detect outlier events.

% The introduction of description complexity and the subsequent memorability measure allow to compare events of various types with different characteristics. As such, it provides a formal and generic framework to discriminate peculiar events and outliers.

% The computation being based on predicates, the memorability score is entirely dependent on their definitions. This can be seen as a caveat or a feature of our approach: it makes the measure subjective while enabling the consideration of user's preferences in the definition and choice of predicates. This consideration can be the topic of future research and open up potential applications.
